### 2019-10-09 thesis/txt/0-introduction.tex
writing on models chapter, putting pictures into figures, update label names

This thesis first presents an overview over the related research, spanning from CI research, build log analysis and augmentation, and common \todo{production?} log processing over information extraction and retrieval techniques to program synthesis by examples.
Next, chapter \ref{models} present our models for information extraction techniques, use case scenarios for such techniques and extractable information from CI build logs. It also introduces the three techniques, we focus on during our research evaluation, as well as their theoretical foundations.
Chapter \ref{data-set} describes the creation of our \emph{The Failing Build Logs Data Set} collected from failed Travis CI build logs, and the labeling and validation process.
Following, Chapter \ref{implementation} our implementation of the three chosen information extraction techniques and the usage of our unified information extraction tool.
This tool is also used for the evaluation study of our models described in Chapter \ref{evaluation}. There we also describe the resulting instantiation of our recommendation model for regex program synthesis, text similarity and keyword search.
Lastly, we conclude and give an overview of further research opportunities in Chapter \ref{conclusion}.
\end{document}

>>>>

This thesis first presents an overview over the related research, spanning from CI research, build log analysis and augmentation, and common \todo{production?} log processing over information extraction and retrieval techniques to program synthesis by examples.
Next, chapter \ref{sec:models} present our models for information extraction techniques, use case scenarios for such techniques and extractable information from CI build logs. It also introduces the three techniques, we focus on during our research evaluation, as well as their theoretical foundations.
Chapter \ref{sec:data-set} describes the creation of our \emph{The Failing Build Logs Data Set} collected from failed Travis CI build logs, and the labeling and validation process.
Following, Chapter \ref{sec:implementation} our implementation of the three chosen information extraction techniques and the usage of our unified information extraction tool.
This tool is also used for the evaluation study of our models described in Chapter \ref{sec:study}. There we also describe the resulting instantiation of our recommendation model for regex program synthesis, text similarity and keyword search.
Lastly, we conclude and give an overview of further research opportunities in Chapter \ref{sec:conclusion}.
\end{document}

### 2019-10-09 thesis/txt/2-models.tex
writing on models chapter, putting pictures into figures, update label names

\chapter{Information Extraction Techniques}
\label{models}
\includegraphics[page=1, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/RetreatSeptemberSlides.pdf}
Many different information extraction techniques ... this section describes the concepitonal / theoretical results of our thesis ... 
We create a model to characterize BLIE techniques so we can discuss their similarities and differences more clearly. 
Additionally we present a complimentary model for IE tasks, to describe use cases for BLIE techniques in a structured way.
From our analysis of build logs we share our notion of the extractable information within a build log.

\section{Characteristics of a Buildlog}
various tools writing in there, each own internal syntax, no common formatting agreements across tools, Travis Output additional newlines / coloring

@@

\section{Our Model for Information Extraction Techniques}
\includegraphics[page=3, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/flow-of-research.pdf}
\plan{model = making a picture with boxes to visualize what we are talking about}
A build log information extraction technique (BLIE technique) consumes a build log (text file) and produces a certain output (text). \todo{make picture, output can be contained in log or not, can be continuous substring or selected lines etc.}.
\begin{itemize}
  \item{Target}{Extractable Information} each extraction technique \todo{instantiation!} targets a specific information
  \item{Setup Overhead}{Time}
  \item{Performance}{Performance} split into learning performance
\end{itemize}

@@

\section{Our Model for Information Extraction Tasks}
\includegraphics[page=4, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/flow-of-research.pdf}
aims to describe an information extraction task of a developer or a researcher.
\begin{itemize}
  \item \todo{describe all classes}
  \item{Setup Overhead}{Time}
  \item{Performance}{Performance} split into learning performance
\end{itemize}
\todo{give 2-3 concrete example instantiations here}

\section{Our Model for Extractable Information in Build Logs}
\includegraphics[page=2, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/flow-of-research.pdf}
during our inital exploration \& data collection for the Failing Build Log Data Set, wanted to get an impression about the information one would want to extract from a build log
\begin{itemize}
  \item \todo{describe all classes}
  \item{Setup Overhead}{Time}
  \item{Performance}{Performance} split into learning performance
\end{itemize}
\todo{differentiate Travis CI specific things, explain high variability for instantiations}
\todo{give at least one concrete example instantiation, best from a log in the data set, not too long so maybe it goes in the abstract?, picture/graphic with marked parts here?}

\section{PROSE Program Synthesis}
\includegraphics[page=2, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/RetreatSeptemberSlides.pdf}
\todo{explanation of process, the way our tool does it, instantiate model of extraction technique, probably no additional theoretical stuff really needed here?, describe regex programs}

\section{IR Similarity}
\includegraphics[page=3, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/RetreatSeptemberSlides.pdf}
\todo{explanation of process, the way our tool does it, instantiate model of extraction technique, probably no additional theoretical stuff really needed here?}

\section{Keyword search}
\includegraphics[page=4, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/RetreatSeptemberSlides.pdf}
\todo{explanation of process, the way our tool does it, instantiate model of extraction technique, probably no additional theoretical stuff really needed here?}

>>>>

\chapter{Information Extraction Techniques}
\label{sec:models}
%\includegraphics[page=1, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/RetreatSeptemberSlides.pdf}
Many different information extraction techniques ... this section describes the conceptional / theoretical results of our thesis ...
We create a model to characterize BLIE techniques so we can discuss their similarities and differences more clearly.
Additionally we present a complimentary model for IE tasks, to describe use cases for BLIE techniques in a structured way.
From our analysis of build logs we share our notion of the extractable information within a build log.

\section{Characteristics of a Build Log}
The idea of CI is, to integrate new software changes fast and often to catch errors early. After making a change, the developer commits and pushes it to a shared source code repository. Companies often use a specific CI server, e.g. Travis CI, linked to their source code repository. A CI build might then be triggered by a push on specific branches or after a pull request was created. When such a CI build is triggered, it typically runs through these stages:
\begin{itemize}
	\item Pulling in the new, changed version of the source code.
	\item Running static analysis tools~\cite{zampetti2017open}.
	\item Building the software, i.e. compiling and packaging it~\cite{phillips2014understanding}.
	\item Running automated tests.
\end{itemize}
However these are only \emph{typical} stages and there is a high variability in the CI build processes of different software projects \mention{citation needed, zampetti2017open?}. Some smaller projects might just use CI to make sure their code compiles as a minimal check before reviewing a pull request, other projects might have various stages of extensive automated testing.

Independent from the actual composition of the CI build process, most software tools used within a build will write out log messages to the console \mention{citation needed, paper about usages of console.log?}. Often, the console output is the only means to for a build tool to communicate with its users. Tools print progress updates, error and warning messages on the console.

The structure of this output is chosen by every tool themselves. Many have implicit or explicit structuring rules, some adhere to predefined standards \mention{phpunit, ruby, xuint output?}. However, when anlayzing build logs we might not have access to the exact build configuration (which tools are used in which order) or might not have access to a useable definition of the output sturcture of a specific tool. Therefore we describe build logs as semi-structured.

Semi-structured for build logs means for us that the log text is:
\begin{itemize}
	\item[implicitly stuructured], we do not have access to explicit structuring elements or an explicit structure description
	\item[irregular], the structure changes without notice \todo{we saw this in our data collection, random newlines}
\end{itemize}

@@

\section{Our Model for Extractable Information in Build Logs}
\begin{figure}[h]
  \centering
\includegraphics[page=2, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/flow-of-research.pdf}
  \caption{Information extractable from build logs}
  \label{fig:build-log-information-draft}
\end{figure}
Continuous integration build logs contain a huge amount of information about the various stages in the CI build they correspond to. In this section we present our model to more precisely describe our notion of a information extractable from a CI build log.

A \emph{build log information} for us has:
\begin{itemize}
	\item a \emph{textual representation} within the build log, more specifically a substring of the log string
\end{itemize}
With this build information items can be hierachically ordered by their textual representations containing each other.

\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth]{img/mt-graphics-BuildLogInformation.pdf}
	\caption{Build Log Information}
	\label{fig:build-log-information}
\end{figure}
During our inital exploration \& data collection for the \emph{Failing Build Log Data Set}, wanted to get an impression about the information one would want to extract from a build log. Within the Travis CI build logs we found various different informations possibly interesting to be extracted
\begin{itemize}
	\item[Travis Build Step] A Travis CI build consists of several build steps defined within the Travis CI configuration language. Within the build log each of these steps are framed by \lstinline{travis_fold:start:<build step name>} and \lstinline{travis_fold:end:<build step name>}.
	      A build step contains
	      \begin{itemize}
		      \item[Build Step Name] The string Travis uses to identify the step within start \& end statements.
		      \item[Build Step Output] The output generated during the build step. The textual representation of the \texttt{Build Step Output} is the substring between the start \& end statements.
	      \end{itemize}
	\item[Travis Timing] Travis can measure the time of specified sections of the build process. It represents those by \lstinline{travis_time:start:<timing section id>} and \lstinline{travis_time:end:<timing section id>:start=<start time>,finish=<finish time>,duration=<duration>}
\end{itemize}

\todo{differentiate Travis CI specific things, explain high variability for instantiations}
\todo{give at least one concrete example instantiation, best from a log in the data set, not too long so maybe it goes in the abstract?, picture/graphic with marked parts here?}

\section{Our Model for Information Extraction Techniques}
\begin{figure}[h]
	\centering
	\includegraphics[page=3, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/flow-of-research.pdf}
	\caption{Model for an information extraction technique}
	\label{fig:model-ie-technique}
\end{figure}
\plan{model = making a picture with boxes to visualize what we are talking about}
A build log information extraction technique (BLIE technique) consumes a build log (text file) and produces a certain output (text). \todo{make picture, output can be contained in log or not, can be continuous substring or selected lines etc.}.
\begin{itemize}
	\item{Target}{Extractable Information} each extraction technique \todo{instantiation!} targets a specific information
	\item{Setup Overhead}{Time}
	\item{Performance}{Performance} split into learning performance
\end{itemize}

@@

\section{Our Model for Information Extraction Tasks}
\begin{figure}[h]
	\centering
	\includegraphics[page=4, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/flow-of-research.pdf}
	\caption{Model for information extraction task}
	\label{fig:modelt-ie-task}
\end{figure}
aims to describe an information extraction task of a developer or a researcher.
\begin{itemize}
	\item \todo{describe all classes}
	\item{Setup Overhead}{Time}
	\item{Performance}{Performance} split into learning performance
\end{itemize}
\todo{give 2-3 concrete example instantiations here}

\section{PROSE Program Synthesis}
\begin{figure}[h]
  \centering
\includegraphics[page=2, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/RetreatSeptemberSlides.pdf}
  \caption{Synthesizing regular expression programs by example using PROSE}
  \label{fig:prose-explanation}
\end{figure}
\todo{explanation of process, the way our tool does it, instantiate model of extraction technique, probably no additional theoretical stuff really needed here?, describe regex programs}

\section{Text Similarity}
\begin{figure}[h]
  \centering
\includegraphics[page=3, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/RetreatSeptemberSlides.pdf}
  \caption{Extracting information using text similarity}
  \label{fig:text-similarity-explanation}
\end{figure}
\todo{explanation of process, the way our tool does it, instantiate model of extraction technique, probably no additional theoretical stuff really needed here?}

\section{Keyword search}
\begin{figure}[h]
  \centering
\includegraphics[page=4, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/RetreatSeptemberSlides.pdf}
  \caption{Extracting information using keyword search}
  \label{fig:keword-search-explanation}
\end{figure}
\todo{explanation of process, the way our tool does it, instantiate model of extraction technique, probably no additional theoretical stuff really needed here?}

### 2019-10-09 thesis/txt/3-data-set.tex
writing on models chapter, putting pictures into figures, update label names

\label{data-set}
\section{Motivation}
originally: log collection to get impression of build logs from various projects, what possibly extractable information they contain and how strutured or unstructured they are.
Then labeling of one build log information, namely the reason the build failed, as well as keywords and stuructural category to use in our evaluation and support our impressions about suitability of the techniques with quantitative results

@@

\section{Labeling}
For our study, described in Chapter \ref{study}, we want to investigate the extraction capabilities of three techniques. Therefore, we need a data set of in/output examples for one \texttt{BuildLogInformation}. Furthermore, Keyword search is not configured by such in/output, but search keywords surrounding the desired output. Finally, this section describes our notion of structural categories, which we use to quantify our assumptions on the needed uniformness of in/output examples.

>>>>

\label{sec:data-set}
\section{Motivation}
originally: log collection to get impression of build logs from various projects, what possibly extractable information they contain and how strutured or unstructured they are.
Then labeling of one build log information, namely the reason the build failed, as well as keywords and stuructural category to use in our evaluation and support our impressions about suitability of the techniques with quantitative results

\section{Related Data Sets}
\todo{kind of related work for data set? the two mortiz sent, more?}

@@

\section{Labeling}
For our study, described in Chapter \ref{sec:study}, we want to investigate the extraction capabilities of three techniques. Therefore, we need a data set of in/output examples for one \texttt{BuildLogInformation}. Furthermore, Keyword search is not configured by such in/output, but search keywords surrounding the desired output. Finally, this section describes our notion of structural categories, which we use to quantify our assumptions on the needed uniformness of in/output examples.

### 2019-10-09 thesis/txt/4-tool.tex
writing on models chapter, putting pictures into figures, update label names

\chapter{Our Information Extraction Tool}
\label{implementation}
describe implemantation of Tool
\includegraphics[page=7, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/flow-of-research.pdf}

\section{Common Interface} 
unification tool, reference CLI somehow nicely?,

>>>>

\chapter{Our Information Extraction Tool}
\label{sec:implementation}
describe implemantation of Tool
\begin{figure}[h]
	\centering
	\includegraphics[page=7, width=\textwidth, trim={0.5cm 0.5cm 0.5cm 0.5cm}, clip]{img/flow-of-research.pdf}
	\caption{Our tool unifying different information extraction techniques for build logs}
	\label{fig:tool}
\end{figure}

\section{Common Interface}
\begin{lstlisting}
Usage: ruby run-extraction.rb -a analyze -t <technique: ir, pbe, random> -e <example_set> -p <path_to_file_to_analyze>
       ruby run-extraction.rb -a evaluate -t <technique: ir, pbe, keyword, random> -e <example_set> -s <selection_technique> -l <step_count_for_learning> -c <test_count>
       ruby run-extraction.rb -a analyze -t keyword -k <keyword, keyword, ...> -p <path_to_file_to_analyze>
       ruby run-extraction.rb -a analyze -t regex -r <regex to match extraction> -p <path_to_file_to_analyze>

Specific options:
    -a, --action ACTION              Either run an extraction for a example set ('analyze') or run the whole evaluation of it ('evaluate')
    -t, --technique TECHNIQUE        The technique used for creating the extraction program (pbe, ir, keyword, random)
    -e, --example-set EXAMPLE_SET    The filename of the example set to use
    -p, --path PATH                  The path to the file to be analyzed relative to the 'tool/samples' folder
    -s, --selection SELECTION        The example sequence selection technique to use for evaluation (chronological, random, manual (= like defined in file))
    -l, --learning-step-count COUNT  How many steps with increasing example set size to do during evaluation
    -c, --test-count COUNT           How many test files to evaluate the generated program in each learning step of the evaluation
    -v, --verbose                    Print additional interesting output apart from only the extraction output
    -k, --keywords X,Y,Z             Keywords too filter lines for during keyword search based extraction
    -r, --regex REGEX                Regex to match on build log file content in regex based extraction

Common options:
    -h, --help                       Show this message
\end{lstlisting}
unification tool, reference CLI somehow nicely?,
